---
title: "MFQ"
author: "Nisha Bharat"
date: "February 14, 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Load in library

```{r include=FALSE}
library(car)
library(ggplot2)
library(MASS)
library(Hmisc)
library(tidyr)
library(tidyverse)
library(dplyr)
library(plyr)
library(summarytools)
library(psych)
```

First, you should download your survey data from Redcap in the form of a csv file. Instructions are on the Github readme. 

Here is a pdf about the SCARED for reference: https://sspediatricassociates.com/Forms-and-Policies/Forms/Behavioral,-Mental-Health-Assessment-Forms/SCARED-form-Parent-and-Child-version.aspx

You should find a reference for your specific measure. To do this, you should first look at the Codebook (GoogleDrive -> Undergrad resources -> Codebook Materials). In there are multiple GoogleDocs that include links to Box that contain material-relevant pdfs. The GoogleDocs also contain citations for the original measure.

If you are confused about any functions/commands used below, type question mark then the name of the command, e.g. ?library. This will provide a description and a few examples about how that function is used.

# Pre-Analysis Cleaning for Two Year Data

## Load in Two Year data

Change the below syntax to fit your computer and where you put your data file, like Desktop or Documents. Replace scared_twoyear with whatever you want to call your dataset, e.g. data or nameofmeasure, just make sure you'll be able to tell that it's the two year data. 

```{r}
mfq_twoyear <- read.csv("C:/Users/nishab/Desktop/mfq_twoyear.csv")
```

## Remove empty rows

Replace scared_twoyear with whatever you want to call your dataset and replace scared1_1_2yr with whatever the first item of the dataset. The syntax removes all rows where there is NA for the first item, so theoretically, this should remove everyone who hasn't completed the Qualtrics survey first. This isn't perfect, since some people may have just happened to skip the first item, but this will do for preliminary statistics. 

```{r}
mfq_twoyear <- mfq_twoyear[!is.na(mfq_twoyear$mfqc1_1_2yr),]
```

## Re-scaling variables

This requires a bit of foresight, but eventually you're going to want to compare the two year data with the original data. I know that the original SCARED data follows a 0-1-2 scale, but the two year data uses a 1-2-3 scale, so we're going to want to always alter the two year data to fit the original data. 

There is probably an easier way to recode multiple variables at once, instead of recoding each variable at a time, so let me know if you find an alternative!

```{r}
mfq_twoyear$mfqc1_1_2yr <- car::recode(mfq_twoyear$mfqc1_1_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc1_2_2yr <- car::recode(mfq_twoyear$mfqc1_2_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc1_3_2yr <- car::recode(mfq_twoyear$mfqc1_3_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc1_4_2yr <- car::recode(mfq_twoyear$mfqc1_4_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc1_5_2yr <- car::recode(mfq_twoyear$mfqc1_5_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc1_6_2yr <- car::recode(mfq_twoyear$mfqc1_6_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc2_7_2yr <- car::recode(mfq_twoyear$mfqc2_7_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc2_8_2yr <- car::recode(mfq_twoyear$mfqc2_8_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc2_9_2yr <- car::recode(mfq_twoyear$mfqc2_9_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc2_10_2yr <- car::recode(mfq_twoyear$mfqc2_10_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc2_11_2yr <- car::recode(mfq_twoyear$mfqc2_11_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc2_12_2yr <- car::recode(mfq_twoyear$mfqc2_12_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc3_13_2yr <- car::recode(mfq_twoyear$mfqc3_13_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc3_14_2yr <- car::recode(mfq_twoyear$mfqc3_14_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc3_15_2yr <- car::recode(mfq_twoyear$mfqc3_15_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc4_20_2yr <- car::recode(mfq_twoyear$mfqc4_20_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc4_21_2yr <- car::recode(mfq_twoyear$mfqc4_21_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc4_22_2yr <- car::recode(mfq_twoyear$mfqc4_22_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc4_23_2yr <- car::recode(mfq_twoyear$mfqc4_23_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc4_24_2yr <- car::recode(mfq_twoyear$mfqc4_24_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc5_25_2yr <- car::recode(mfq_twoyear$mfqc5_25_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc5_26_2yr <- car::recode(mfq_twoyear$mfqc5_26_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc5_27_2yr <- car::recode(mfq_twoyear$mfqc5_27_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc5_28_2yr <- car::recode(mfq_twoyear$mfqc5_28_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc5_29_2yr <- car::recode(mfq_twoyear$mfqc5_29_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc6_30_2yr <- car::recode(mfq_twoyear$mfqc6_30_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc6_31_2yr <- car::recode(mfq_twoyear$mfqc6_31_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc6_32_2yr <- car::recode(mfq_twoyear$mfqc6_32_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc6_33_2yr <- car::recode(mfq_twoyear$mfqc6_33_2yr, "1=0; 2=1; 3=2")
mfq_twoyear$mfqc6_34_2yr <- car::recode(mfq_twoyear$mfqc6_34_2yr, "1=0; 2=1; 3=2")

```

## Make a sum and a mean variable

The below syntax creates two variables and adds them back to the dataset: the mean of all the items and the sum of all the items.

```{r}
names(mfq_twoyear) #This prints the dataset's variables names so, which makes it so I don't totally have to write out each individual variable name

mfq_vars_dummy <- cbind(mfq_twoyear$mfqc1_1_2yr, mfq_twoyear$mfqc1_2_2yr, mfq_twoyear$mfqc1_3_2yr, mfq_twoyear$mfqc1_4_2yr, mfq_twoyear$mfqc1_5_2yr, mfq_twoyear$mfqc1_6_2yr, mfq_twoyear$mfqc2_7_2yr, mfq_twoyear$mfqc2_8_2yr, mfq_twoyear$mfqc2_9_2yr, mfq_twoyear$mfqc2_10_2yr, mfq_twoyear$mfqc2_11_2yr, mfq_twoyear$mfqc2_12_2yr, mfq_twoyear$mfqc3_13_2yr, mfq_twoyear$mfqc3_14_2yr, mfq_twoyear$mfqc3_15_2yr, mfq_twoyear$mfqc4_20_2yr, mfq_twoyear$mfqc4_21_2yr, mfq_twoyear$mfqc4_22_2yr, mfq_twoyear$mfqc4_23_2yr, mfq_twoyear$mfqc4_24_2yr, mfq_twoyear$mfqc5_25_2yr, mfq_twoyear$mfqc5_26_2yr, mfq_twoyear$mfqc5_27_2yr, mfq_twoyear$mfqc5_28_2yr, mfq_twoyear$mfqc5_29_2yr, mfq_twoyear$mfqc6_30_2yr, mfq_twoyear$mfqc6_31_2yr, mfq_twoyear$mfqc6_32_2yr, mfq_twoyear$mfqc6_33_2yr, mfq_twoyear$mfqc6_34_2yr)
```

```{r}
mfq_twoyear$mfq_total_mean <- rowMeans(mfq_vars_dummy, na.rm=TRUE)
```

```{r}
mfq_twoyear$mfq_total_sum <- rowSums(mfq_vars_dummy, na.rm=TRUE)
```

```{r}
rm(mfq_vars_dummy)
```

Depending on your measure, you might want to have a mean and a sum variable for each subscale. You would use rowSums() and rowMeans() respectively. The above chunk only makes a sum variable for the Panic Disorder subscale.

Whoever is doing the SCARED dataset should do the other subscales here.

# Descriptive statistics for two year data

Now you can start to descriptive statistics. Below is a list of things that you might want to calculate for all the items. Make sure to be careful to not run statistics on the *entire* dataset, as the entire dataset includes id, scared_youth_paper_2yr, and scared_youth_complete, and those aren't meaningful variables.

* Sum
* Mean
* Median
* Std dev
* Std error
* Min/max/range
* IQR
* Kurtosis
* Skewness

Luckily, instead of writing out syntax for each statistic (e.g. mean(), sd(), etc.), there are funcitons that will write out a nice descriptive statistics table.

```{r, results='asis'}
summarytools::descr(mfq_twoyear, style = "rmarkdown", na.rm=TRUE)
```

You can also visualize the data.

```{r}
hist(mfq_twoyear$mfq_total_sum)
hist(mfq_twoyear$mfq_total_mean)
```

Feel free to try out any stats you want or to visually improve the look of your tables or visualizations. Once you've done a bunch of statistics and you want to see how it all looks, click the *Knit* button at the top (next to the ball of yarn) and choose "Knit to HTML". Knitting or clicking "Run All" in the *Run* drop-down menu is a good test to see if there are any errors in your code. 

```{r}
mfq_twoyear$mfq_youth_paper_2yr <- NULL
mfq_twoyear$mfqyouth2yr_complete <- NULL
```


```{r}
which(mfq_twoyear == -99, arr.ind=TRUE)
```


# Comparing Wave 15 and Two Year data

First, follow the instructions on the Github readme to download the original Wave 15 data for your measure. Once you've done that, read in your data.

## Load in Wave 15 data
```{r}
mfq_original <- read.csv("C:/Users/nishab/Desktop/mfq_original.csv")
names(mfq_original)[names(mfq_original) == "ï..id"] <- "id"
```
#Don't ask why the id variable looks funny, it just does that sometimes when exporting/importing.

Luckily, graduate students have polished most of the original Wave 15 data, so things like total sum/mean variables and subscales already have their own columns. They may go by different names, so be sure to look at the original dataset and its variable names to see what's there.

```{r}
View(mfq_original)

```

## Merge the original and two year datasets

Merging the datasets drops all the IDs in scared_original that aren't present in scared_twoyear. Since we've combined the datasets into a new dataset ("scared"), there is no need to the keep the old datasets (and any objects associated with either of those datasets) in our environment. Additionally, as long as you don't move your data files and you keep all the code above, you can get them right back where they were if you realize that you need them back. 

```{r}
mfq <- merge(mfq_original, mfq_twoyear, by="id")
```

```{r}
rm(mfq_twoyear)
rm(mfq_original)
```

## Add demographic data

There is a file called "demographic.csv" and it is located in the "Data Cleaning Protocol" on the Fragile Families server (it's at the very bottom). Copy it and paste the copy to your desktop. At the moment, this file only includes gender ("F" or "M") and age (as MM/DD/YYYY). Race, ethnicity, SES, and other variables might be added to demographic.csv later. 

```{r}
demographic <- read.csv("C:/Users/nishab/Desktop/demographic.csv")
mfq <- merge(mfq, demographic, by="id")
rm(demographic)
```

## Cleaning/managing the combined dataset

Just by looking at my dataset, you can tell that some people decided to entirely skip the SCARED survey during their Wave 15 visit. I can tell this because there's a long row of NA's or 88's. For SCARED, these people include 10134 and 10219, which are rows 51 and 89 respectively. It might be best to remove them. We'll use the same syntax that we used to drop empty rows at the beginning.

```{r}
scared <- scared[-c(51, 89), ]
```

Also, whenever a value is missing or not applicable, we've been coding it as 88 and -99. We should take all the values that equal 88 or -99 and make them NA.

```{r}
mfq[mfq=="88"] <- NA
```

## Descriptive statistics on the combined dataset

Now we have a dataset that includes the original and two year data, so we can start answering questions: have participants' anxiety gotten worse over two years? Gotten better? Maybe worse with panic disorder but not with social anxiety? 

Feel free to add age or gender to your analyses, but just be careful with how the data is formatted. You might want to re-code gender so instead of it being M/F, it would be 0=Male and 1=Female. Also for age, you might to want make TeenBirthday a numeric variable by subtracting it from today's date.

```{r Basic stuff}
mean(scared$SCARED_C_Sum)
mean(scared$scared_total_sum)

sd(scared$SCARED_C_Sum)
sd(scared$scared_total_sum)

hist(scared$SCARED_C_Sum)
hist(scared$scared_total_sum)
```

```{r}
cor.test(scared$SCARED_C_Sum, scared$scared_total_sum) #Is there a correlation between the old and new scores?
```

```{r}
t.test(scared$SCARED_C_Sum, scared$scared_total_sum) #Is there a significant difference between these two scores?
```

```{r}
plot(scared$SCARED_C_Sum, scared$scared_total_sum)
```
